---
title: "Ph1 analysis"
author: "Matt Sant-Miller"
date: "10/10/2021"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(magrittr)
```

## Step 1: Clean data

To begin, we perform basic cleaning on the data using the raw data available at 
[this](https://docs.google.com/spreadsheets/d/1pS3TBai5-lu-xYGRImVZ-AZ250Q8lvLwJ7c0hw2EEA4/edit#gid=949930203)
Google Sheet.

```{r, echo = TRUE, message = FALSE, warning = FALSE}
# Define function to load data from Google Sheet--------------------------------
GetCleanPh1Data <- function(input_sheet){
  googlesheets4::read_sheet(
    ss = "https://docs.google.com/spreadsheets/d/1pS3TBai5-lu-xYGRImVZ-AZ250Q8lvLwJ7c0hw2EEA4",
    sheet = input_sheet
  ) %>%
    dplyr::rename_all(.funs = stringr::str_to_lower) %>%
    janitor::clean_names()
}

# Load and clean variable definitions-------------------------------------------
ph1_definitions <-
  GetCleanPh1Data(input_sheet = "variable_definitions") %>%
  dplyr::filter(status == "Active")

# Create variable subsets-------------------------------------------------------
ph1_variables <- tibble::lst()
for(class_i in c("outcome","predictor")){
  for(status_i in c("clean","raw")){
    ph1_variables[[base::paste(class_i, status_i, sep = "_")]] <-
      ph1_definitions %>%
      dplyr::filter(class == stringr::str_to_sentence(class_i)) %>%
      purrr::pluck(base::paste(status_i, "name", sep = "_"))
  }
}
for(type_i in c("categorical","numeric")){
  for(status_i in c("clean","raw")){
    ph1_variables[[base::paste(type_i, status_i, sep = "_")]] <- 
      ph1_definitions %>%
      dplyr::filter(type == stringr::str_to_sentence(type_i)) %>%
      purrr::pluck(base::paste(status_i, "name", sep = "_"))
    ph1_variables[[base::paste("predictor", type_i, status_i, sep = "_")]] <- 
      ph1_definitions %>%
      dplyr::filter(class == "Predictor", type == stringr::str_to_sentence(type_i)) %>%
      purrr::pluck(base::paste(status_i, "name", sep = "_"))
  }
}
for(status_i in c("clean","raw")){
  ph1_variables[[base::paste("log", status_i, sep = "_")]] <-
    ph1_definitions %>%
    dplyr::filter(transformation == "Log") %>%
    purrr::pluck(base::paste(status_i, "name", sep = "_"))
}

# Load and clean Ph1 data-------------------------------------------------------
ph1_data <-
  GetCleanPh1Data(input_sheet = "data") %>%
  # Clean categorical data
  dplyr::mutate(
    prev_season_vacc_status = dplyr::case_when(
      prev_season_vacc_status == "UKNOWN" | 
        prev_season_vacc_status == "UNKNOWN" |
        prev_season_vacc_status == "UnkNOWN" ~ NA_character_,
      T ~ prev_season_vacc_status
    )
  ) %>% dplyr::mutate_at(
    .vars = ph1_variables[["categorical_raw"]],
    .funs = base::as.factor
  ) %>%
  # Clean numeric data
  dplyr::mutate_at(
    .vars = ph1_variables[["log_raw"]],
    .funs = tibble::lst(log = ~dplyr::case_when(
      . == 0 ~ 0,
      T ~ base::log(.)
    ))
  ) %>%
  dplyr::mutate(quantity_minus_efficacy = ph1_fc_tot_igg_log - x1_kd_fc_ph1_ha1_log) %>%
  dplyr::mutate(ph1_delta_tot_igg = ph1_d14_tot_igg - ph1_d0_tot_igg) %>%
  dplyr::mutate(one_over_ph1_d0_tot_igg = 1/ph1_d14_tot_igg)

```

## Step 2: Analyze data

To explore the relationship between antibody counts/affinity and CD4 T-Cell counts,
we run a multivariate regression, predicting fold change in CD4 ELISpot as a function
of baseline antibody count (total IgG), baseline antibody affinity (1/kd), and the
interaction between the two metrics. The regression coefficients on each of these
terms are found to be statistically significant with 95 percent confidence.
We visualize these predictive relationships using a three-dimensional plot, in which
baseline antibody count is on the X-axis, baseline antibody affinity is on the Y-axis
and the predicted mean of fold change is plotted on the vertical Z-axis.
```{r, include = TRUE}
model <- stats::lm(
  formula = ph1_fc_elispot_log ~ 
    x1_kd_d0_ph1_ha1_log + ph1_d0_tot_igg_log +
    x1_kd_d0_ph1_ha1_log * ph1_d0_tot_igg_log,
  data = ph1_data
)
base::summary(model)
ph1_new_data <- 
  tidyr::expand_grid(
    x1_kd_d0_ph1_ha1_log = base::seq(
      from = 3,
      to = 8,
      by = 0.1
    ),
    ph1_d0_tot_igg_log = base::seq(
      from = 5,
      to = 10,
      by = 0.1
    )
  )
model_predictions <-
  dplyr::bind_cols(
    ph1_new_data,
    ph1_fc_elispot_log_pred = stats::predict(
      object = model,
      newdata = ph1_new_data
    )
  )

plot <-
  plotly::plot_ly() %>%
  plotly::add_trace(data = model_predictions,
                    x = ~x1_kd_d0_ph1_ha1_log,
                    y = ~ph1_d0_tot_igg_log,
                    z = ~ph1_fc_elispot_log_pred,
                    type = "mesh3d")
plot
```
